Epidemic Tracking and Forecasting: Lessons Learned from a Tumultuous Year

### 2021-11-29 cover/cover.Rmd
Make eval=TRUE
Fix latex typo

```{r get-data, eval = FALSE}
sources <- c("jhu-csse",
             "doctor-visits",
             "fb-survey",
             "google-symptoms")

@@

```{r load-all}
load(file = "all.rda")
```

@@

>>>>

```{r get-data}
sources <- c("jhu-csse",
             "doctor-visits",
             "fb-survey",
             "google-symptoms")

@@

```{r load-all, eval = FALSE}
load(file = "all.rda")
```

@@

```{r trend-plots-no-y-labels}
for (i in 1:length(x)) {
  g[[i]] <- x[[i]] %>%
    full_join(y, by = c("geo_value", "time_value")) %>%
    mutate(geo_num = as.integer(as.factor(geo_value))) %>%
    group_by(geo_value) %>%
    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) +
             delta * (geo_num - 1)) %>%
    ungroup() %>%
    ggplot(aes(x = time_value, y = value, group = geo_value)) +
    labs(y = names[i]) + coord_cartesian(xlim = xlim, ylim = ylim) +
    geom_line(alpha = 0.5, color = colors[i], size = line_width) +
    theme_void() +
    theme(panel.background = element_rect(fill = "black"))
}

gg <- do.call(gridExtra::arrangeGrob, list(grobs = g, ncol = 1))
ggsave(file = "cover-no-labs.pdf", gg, height = 22.5, width = 21, units = "cm")
ggsave(file = "cover-no-labs.png", gg, height = 15.9, width = 15, units = "cm")
```

### 2021-11-29 perspective/paper.tex
Fix latex typo

Epidemic forecasting has garnered increasing interest in the last decade,
nurtured and scaffolded by various forecasting challenges organized by groups
within the U.S. federal government, including the CDC \citep{Biggerstaff:2016,
  Biggerstaff:2018, Reich:2019}, OSTP \citep{Johansson:2019}, DARPA
\citep{DelValle:2018}, and elsewhere \citep{Ajelli:2018, Viboud:2018}.  In 2017,
after several years of experimentation with flu forecasting in academic groups,
the CDC decided to incorporate influenza forecasting into its normal operations,
including weekly public communications citep{FluSight} and briefing to
higher-ups.  To provide more reliable infrastructure and support for its
forecasting needs, the CDC in 2019 designated two national Centers of Excellence
for Influenza Forecasting, one at the University of Massachusetts at
Amherst\footnote{\url{https://reichlab.io/people}} and one at Carnegie Mellon
University\footnote{\url{https://delphi.cmu.edu/about/center-of-excellence/}}.

@@

In the nearly 10 years of government-organized epidemic forecasting exercises in
the U.S., efforts were focused on modeling the natural history and likely
evolution of the pathogen, with adaptation of human behavior playing a secondary
role (if any role at all).  The pandemic demonstrated that our forecasting
models must pay closer attention to reactive human behavior, even more so if we
are to consider interventions.  Unfortunately, many highly relevant aspects of
human behavior are not measured by publicly available data streams, like
compliance with policies and recommendations (with perhaps surveys and mobility
reports providing our best glimpse into these hard-to-observe aspects of
behavior \citep{Bilinski:2021}), and even the very process of promulgating these 
policies and recommendations.  Furthermore, even if we had these data in hand,
incorporating their effects will require significant and new cognitive and
behavioral modeling, with uncertain success.  The tragic breakdown and
fragmentation of trust in governments, public health officials, and healthcare
professionals are perhaps the hardest factors to measure and model, yet they
played an undeniable role in the progression of the current pandemic in the
U.S. and other countries.

>>>>

Epidemic forecasting has garnered increasing interest in the last decade,
nurtured and scaffolded by various forecasting challenges organized by groups
within the U.S. federal government, including the CDC \citep{Biggerstaff:2016,
  Biggerstaff:2018, Reich:2019}, OSTP \citep{Johansson:2019}, DARPA
\citep{DelValle:2018}, and elsewhere \citep{Ajelli:2018, Viboud:2018}.  In 2017,
after several years of experimentation with flu forecasting in academic groups,
the CDC decided to incorporate influenza forecasting into its normal operations,
including weekly public communications \citep{FluSight} and briefing to
higher-ups.  To provide more reliable infrastructure and support for its
forecasting needs, the CDC in 2019 designated two national Centers of Excellence
for Influenza Forecasting, one at the University of Massachusetts at
Amherst\footnote{\url{https://reichlab.io/people}} and one at Carnegie Mellon
University\footnote{\url{https://delphi.cmu.edu/about/center-of-excellence/}}.

@@

In the nearly 10 years of government-organized epidemic forecasting exercises in
the U.S., efforts were focused on modeling the natural history and likely
evolution of the pathogen, with adaptation of human behavior playing a secondary
role (if any role at all).  The pandemic demonstrated that our forecasting
models must pay closer attention to reactive human behavior, even more so if we
are to consider interventions.  Unfortunately, many highly relevant aspects of
human behavior, like compliance with policies and recommendations, are not
measured by publicly available data streams (with perhaps surveys and mobility  
reports providing our best glimpse into these hard-to-observe aspects of
behavior \citep{Bilinski:2021}).  Furthermore, even if we had these data in
hand, incorporating their effects will require significant and new cognitive and 
behavioral modeling, with uncertain success.  The tragic breakdown and
fragmentation of trust in governments, public health officials, and healthcare
professionals are perhaps the hardest factors to measure and model, yet they
played an undeniable role in the progression of the current pandemic in the
U.S. and other countries.
